---
title: "data_explore"
author: "Bingfeng Xia"
date: "3/29/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 upload data
```{r uploadData}
train = read.csv(file = "train.csv", header = T)
test = read.csv(file = "test.csv", header = T)
```

```{r}
#convert dependent variable 
train$voted = as.numeric(train$voted) 
train$voted[train$voted == 1] = 0
train$voted[train$voted == 2] = 1
```

## 2 data exploration
```{r dataExploration}
summary(train)
```

dbdistance and vccdistance 
gender+cd+hd+age+dbdistance+vccdistance+
party+racename+hsonly+mrrg+chldprsnt+cath+
evang+nonchrst+otherchrst+days.since.reg

gender+cd+hd+age+party+racename+hsonly+mrrg+chldprsnt+cath+
evang+nonchrst+otherchrst+days.since.reg

```{r}
train.na = na.convert.mean(train)
summary(train.na)
```

```{r}
Pidx = sample(nrow(train.na), replace=FALSE)
test.new = train.na[Pidx[1:10000],]
train.new = train.na[Pidx[10000:nrow(train.na)],]
var = c("voted","gender","cd","hd","age","party","racename","hsonly","mrrg",
         "chldprsnt","cath","evang","nonchrst","otherchrst",
         "days.since.reg")
test.new = test.new[,var]
train.new = train.new[,var]
```


```{r}
#basic full logit model
voted.logit0 = glm(voted ~ ., 
                   data = train.new, family = binomial, maxit = 100000)
test.pred = predict(voted.logit0, test.new, type="response")
voted.glm.pred0 = ifelse(test.pred>.5, 1, 0)
res=xtabs(~voted.glm.pred0+test.new$voted)
res
(res[1,1] + res[2,2])/(res[1,1]+res[2,2]+res[1,2]+res[2,1])
```

```{r}
discrepancy<-function(xx,yy) {
  mapped = mapply(function(x,y) x*log(y)+(1-x)*log(1-y),xx,yy)
mean(mapped)}
discrepancy(test.new$voted, test.pred)
```

```{r}
summary(voted.logit0)
```










### colinearity

```{r}
#check collinearity
library(car)
car::vif(voted.logit0)  #imputing 
```

```{r}
par(mfrow = c(2,3))
plot(train.new$cath, train.new$evang)
plot(train.new$cath, train.new$nonchrst)
plot(train.new$cath, train.new$otherchrst)
plot(train.new$evang, train.new$nonchrst)
plot(train.new$evang, train.new$otherchrst)
plot(train.new$nonchrst, train.new$otherchrst)
```


```{r}
#logit model for colinearity
var = c("voted","gender","cd","hd","age","party","racename","hsonly","mrrg",
         "chldprsnt","evang","nonchrst","otherchrst",
         "days.since.reg")
test.coli = test.new[,var]
train.coli = train.new[,var]
voted.logit.coli = glm(voted ~ ., 
                   data = train.coli, family = binomial, maxit = 100000)
test.pred.coli=predict(voted.logit.coli, test.coli, type="response")
voted.glm.pred.coli = ifelse(test.pred.coli>.5, 1, 0)
res=xtabs(~voted.glm.pred.coli+test.coli$voted)
res
(res[1,1] + res[2,2])/(res[1,1]+res[2,2]+res[1,2]+res[2,1])
```
```{r}
discrepancy<-function(xx,yy) {
  mapped = mapply(function(x,y) x*log(y)+(1-x)*log(1-y),xx,yy)
mean(mapped)}
discrepancy(test.coli$voted, test.pred.coli)
```

```{r}
car::vif(voted.logit.coli)  #imputing 
```






### data transformation

#### for paper

```{r}
par(mfrow = c(2,3))
hist(train[,5], main= colnames(train)[5], mgp = c(1,0,0))
hist(train[,10], main= colnames(train)[10], mgp = c(1,0,0))
hist(train[,12], main= colnames(train)[12], mgp = c(1,0,0))
train.trans = train
train.trans$age=log(train$age)
train.trans$chldprsnt=sqrt(train$chldprsnt)
train.trans$hsonly=log(train$hsonly)
hist(train.trans[,5], main= colnames(train.trans)[5], mgp = c(1,0,0))
hist(train.trans[,10], main= colnames(train.trans)[10], mgp = c(1,0,0))
hist(train.trans[,12], main= colnames(train.trans)[12], mgp = c(1,0,0))
```



```{r}
par(mfrow = c(3,3))
for (i in 1:17) {
  if(i == 1 || i == 2 || i == 8 || i == 9) {
    plot(train[,i], main= colnames(train)[i], mgp = c(1,0,0))
  }
  else {
    hist(train[,i], main= colnames(train)[i], mgp = c(1,0,0))
  }
}
```

```{r}
train.trans = train
train.trans$age=log(train$age)
train.trans$dbdistance=log(train$dbdistance)
train.trans$vccdistance=log(train$vccdistance)
train.trans$chldprsnt=sqrt(train$chldprsnt)
train.trans$hsonly=log(train$hsonly)
```


```{r}
par(mfrow = c(3,3))
for (i in 1:17) {
  if(i == 1 || i == 2 || i == 8 || i == 9) {
    plot(train.trans[,i], main= colnames(train.trans)[i], mgp = c(1,0,0))
  }
  else {
    hist(train.trans[,i], main= colnames(train.trans)[i], mgp = c(1,0,0))
  }
}
```


```{r}
train.coli.trans = train.coli
test.coli.trans = test.coli
test.coli.trans$age=log(test.coli$age)
test.coli.trans$chldprsnt=log(test.coli$chldprsnt)
test.coli.trans$hsonly=log(test.coli$hsonly)
```


```{r}
voted.logit.coli.trans = glm(voted ~ ., 
                   data = train.coli.trans, family = binomial, maxit = 100000)

test.pred.coli.trans = predict(voted.logit.coli.trans, test.coli.trans, type="response")
voted.glm.pred.coli.trans = ifelse(test.pred.coli.trans>.5, 1, 0)
res=xtabs(~voted.glm.pred.coli.trans+test.coli.trans$voted)
res
(res[1,1] + res[2,2])/(res[1,1]+res[2,2]+res[1,2]+res[2,1])
```
```{r}
discrepancy<-function(xx,yy) {
  mapped = mapply(function(x,y) x*log(y)+(1-x)*log(1-y),xx,yy)
mean(mapped)}
discrepancy(test.coli.trans$voted, test.pred.coli.trans)
```

```{r}
car::vif(voted.logit.coli.trans)  #imputing 
```










### MyStep model selection
### GLM

```{r}
source("myStep.R")
```

var = c("voted","gender","cd","hd","age","party","racename","hsonly","mrrg",
         "chldprsnt","evang","nonchrst","otherchrst",
         "days.since.reg")

```{r}
train.coli.trans.fac = train.coli.trans
train.coli.trans.fac$gender = as.factor(train.coli.trans.fac$gender)
train.coli.trans.fac$cd = as.factor(train.coli.trans.fac$cd)
train.coli.trans.fac$hd = as.factor(train.coli.trans.fac$hd)
train.coli.trans.fac$party = as.factor(train.coli.trans.fac$party)
train.coli.trans.fac$racename = as.factor(train.coli.trans.fac$racename)

#summary(train.coli.trans.fac$cd)
#nrow(train.coli.trans.fac)
train.coli.trans.fac=train.coli.trans.fac[train.coli.trans.fac$cd != 3.76732727564184,]
summary(train.coli.trans.fac$cd)
#summary(train.coli.trans.fac)
#summary(train.coli.trans.fac$racename)
```
```{r}
test.coli.trans.fac = test.coli.trans
test.coli.trans.fac$gender = as.factor(test.coli.trans.fac$gender)
test.coli.trans.fac$cd = as.factor(test.coli.trans.fac$cd)
test.coli.trans.fac$hd = as.factor(test.coli.trans.fac$hd)
test.coli.trans.fac$party = as.factor(test.coli.trans.fac$party)
test.coli.trans.fac$racename = as.factor(test.coli.trans.fac$racename)
test.coli.trans.fac=train.coli.trans.fac[test.coli.trans.fac$cd != 3.76732727564184,]
summary(test.coli.trans.fac$cd)
```



```{r}
#fit = glm(voted~1,data=train.coli.trans.fac,family="binomial")
#out = myStep(fit,data=train.coli.trans.fac,K=1,test="Chisq",direction="forward",fast=FALSE)
```

```{r}
voted.glm.sel1 = glm(formula = voted ~ hd + age + party + mrrg + 
                       days.since.reg + chldprsnt + gender + racename + 
                       nonchrst + evang + otherchrst + cd, 
                     family = "binomial", data = train.coli.trans.fac)
```

```{r}
voted.logit.coli.trans.13 = glm(formula = voted ~ hd + hsonly + age + party + mrrg + 
                                  days.since.reg + chldprsnt + gender + racename + 
                                  nonchrst + evang + otherchrst + cd, 
                                family = "binomial", data = train.coli.trans.fac)
# final model
voted.glm.sel1 = glm(formula = voted ~ hd + age + party + mrrg + 
                       days.since.reg + chldprsnt + gender + racename + 
                       nonchrst + evang + otherchrst + cd, 
                     family = "binomial", data = train.coli.trans.fac)
anova(voted.logit.coli.trans.13, voted.glm.sel1, test = "Chi")
```


```{r}
voted.glm.sel1.pred=ifelse(predict(voted.glm.sel1, test.coli.trans.fac, type="response")>.5, 1, 0)
res=xtabs(~voted.glm.sel1.pred+test.coli.trans.fac$voted)
res
(res[1,1] + res[2,2])/(res[1,1]+res[2,2]+res[1,2]+res[2,1])
```

```{r}
voted.glm.sel1.pred.response = predict(voted.glm.sel1, test.coli.trans.fac, type="response")
discrepancy<-function(xx,yy) {
  mapped = mapply(function(x,y) x*log(y)+(1-x)*log(1-y),xx,yy)
mean(mapped)}
discrepancy(test.coli.trans.fac$voted, voted.glm.sel1.pred.response)
```

**"hsonly"** deleted

### GAM model selection

```{r}
formula = getGAMformula(voted~hd + age + party + mrrg + days.since.reg + 
    chldprsnt + gender + racename + nonchrst + evang + otherchrst + 
    cd, data=train.coli.trans.fac)
formula 
fit = gam(formula ,data=train.coli.trans.fac, family="binomial")
out = myStep(fit,data=train.coli.trans.fac,
             testMethod="anova",test="Chisq",
             direction="backward",K=1,fast=TRUE)
```


```{r}
# final GAM model
library(mgcv)
voted.gam.sel1 = gam(voted ~ hd + s(age) + party + s(mrrg) + 
                       s(days.since.reg) + s(chldprsnt) + gender + racename + 
                       s(nonchrst) + s(evang) + s(otherchrst) + cd, 
                     family = "binomial", data = train.coli.trans.fac)
anova(voted.glm.sel1, voted.gam.sel1, test = "Chi")
```

```{r}
anova(voted.glm.sel1, voted.gam.sel1, test = "Chi")
```

```{r}
test.pred.gam = predict(voted.gam.sel1, test.coli.trans.fac, type="response")
voted.gam.sel1.pred=ifelse(test.pred.gam>.5, 1, 0)
res=xtabs(~voted.gam.sel1.pred+test.coli.trans.fac$voted)
res
(res[1,1] + res[2,2])/(res[1,1]+res[2,2]+res[1,2]+res[2,1])
```
```{r}
discrepancy<-function(xx,yy) {
  mapped = mapply(function(x,y) x*log(y)+(1-x)*log(1-y),xx,yy)
mean(mapped)}
discrepancy(test.coli.trans.fac$voted, test.pred.gam)
```



## Random Forest Tree

### default r f
```{r}
train.coli.trans.fac.nohd =  train.coli.trans.fac[,!(names(train.coli.trans.fac) 
                                                %in% c("hd"))]
test.coli.trans.fac.nohd =  test.coli.trans.fac[,!(names(train.coli.trans.fac) 
                                                %in% c("hd"))]
```
```{r}
library(randomForest)
voted.rf = randomForest(as.factor(voted)~ age + party + mrrg + 
                       days.since.reg + chldprsnt + gender + racename + 
                       nonchrst + evang + otherchrst + cd, 
                       data=train.coli.trans.fac.nohd, method="class")
print(voted.rf)
```


```{r}
library(randomForest)
voted.rf.prob = randomForest(as.factor(voted)~ age + party + mrrg + 
                       days.since.reg + chldprsnt + gender + racename + 
                       nonchrst + evang + otherchrst + cd, 
                       data=train.coli.trans.fac.nohd, method="prob")
print(voted.rf.prob)
```



```{r}
voted.rf.pred.prob=predict(voted.rf, test.coli.trans.fac.nohd, type="prob")
```

```{r}
#voted.rf.pred.prob[,2]
```

```{r}
voted.rf.pred.prob.pro=ifelse(voted.rf.pred.prob[,2]>.5, 1, 0)
res=xtabs(~voted.rf.pred.prob.pro+test.coli.trans.fac.nohd$voted)
res
(res[1,1] + res[2,2])/(res[1,1]+res[2,2]+res[1,2]+res[2,1])
```


```{r}
discrepancy<-function(xx,yy) {
  mapped = mapply(function(x,y) x*log(y)+(1-x)*log(1-y),xx,yy)
mean(mapped)}
discrepancy(test.coli.trans.fac.nohd$voted, voted.rf.pred.prob[,2])
```


```{r}
voted.rf.pred=predict(voted.rf, test.coli.trans.fac.nohd, type="class")
res=xtabs(~voted.rf.pred+test.coli.trans.fac.nohd$voted)
res
(res[1,1] + res[2,2])/(res[1,1]+res[2,2]+res[1,2]+res[2,1])
```


### select para, haven't been done

```{r}
library(randomForest)
cvr = rfcv(na.roughfix(train.coli.trans.fac.nohd[,!(names(train.coli.trans.fac.nohd) 
                                                %in% c("voted"))]),train.coli.trans.fac.nohd$voted,step=0.9) 
cbind(nvars=cvr$n.var, error.rate=cvr$error.cv)
```



## Cook's distances
## Fitted vs deviance residual

### GLM
```{r q1e}
library(arm)
div.fitted = fitted(voted.glm.sel1)
div.devresid = residuals(voted.glm.sel1,type="deviance")
div.cooks = cooks.distance(voted.glm.sel1)

plot(div.fitted, div.devresid,
  xlab="Fitted values",
  ylab="Deviance residuals",
  pch=19, col="red", cex=1.0,
  main="GLM Fitted vs deviance residual")
abline(h=0,lty=2,col="green")
```

```{r}
plot(div.cooks, type="h", lwd=2,
  xlab="Observation",
  ylab="Cook's distances",
  main="GLM Cook's distances")
abline(h=1,lty=2,col="red")
```


### GAM
```{r q1e}
library(arm)
div.fitted = fitted(voted.gam.sel1)
div.devresid = residuals(voted.gam.sel1,type="deviance")
div.cooks = cooks.distance(voted.gam.sel1)

plot(div.fitted, div.devresid,
  xlab="Fitted values",
  ylab="Deviance residuals",
  pch=19, col="red", cex=1.0,
  main="GAM Fitted vs deviance residual")
abline(h=0,lty=2,col="green")
```

```{r}
plot(div.cooks, type="h", lwd=2,
  xlab="Observation",
  ylab="Cook's distances",
  main="GAM Cook's distances")
abline(h=1,lty=2,col="red")
```


```{r}
library(arm)
library(scales)
gam12.fitted = fitted(voted.gam.sel1)
gam12.devresid = residuals(voted.gam.sel1, type="deviance")
plot(gam12.fitted, gam12.devresid,
  xlab="Fitted counts",
  ylab="Deviance residuals",
  pch=19, col=rgb(red=1, green=0.2, blue=0.5, alpha=0.1), cex=0.5,
  main="GAM Fitted vs deviance residual plot")
abline(h=0,lty=2,col="green")
```




